---
title: "STAT/MATH 495: Problem Set 06"
author: "Brenna Sullivan"
date: "2017-10-17"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(mosaic)
library(broom)
library(knitr)
```


# Collaboration

Please indicate who you collaborated with on this assignment: 



# Setup

Define truth, which again we know for the purposes of this assignment, but in
practice we won't:

* the true function f(x) i.e. the signal
* the true epsilon i.e. the noise, which in this case is Normal$(0, sd=\sigma)$.
Hence the standard deviation $\sigma$ determines the amount of noise.

```{r}
f <- function(x) {
  x^2
}
sigma <- 0.3
```

This is the target point we'll be trying to predict: $(0.95, f(0.95)) = (0.95, 0.95^2) = (0.95, 0.9025)$, Thus, the test set is just `x=0.95`

```{r}
x0 <- 0.95
test_set <- data_frame(x=x0)
```

This function generates a random sample of size $n$; think of this as a "get new
data" function. Random in terms of both:

* (New) the predictor x (uniform on [0,1])
* the amount of noise $\epsilon$

```{r}
generate_sample <- function(f, n, sigma) {
  sample <- data_frame(
    x = runif(n = n, min = 0, max = 1),
    f_x = f(x),
    epsilon = rnorm(n = n, mean = 0, sd = sigma),
    y = f_x + epsilon
  )
  # Recall: We don't observe f(x) and epsilon, just (x, y)
  sample <- sample %>% 
    select(x, y)
  
  return(sample)
}
```

Define

* The number $n$ of observations $(x_i, y_i)$ in each sample. In the handout,
$n=100$ to keep plots uncrowded. Here we boost to $n=500$
* Number of samples of size $n$ to consider

```{r}
n <- 500
n_sample <- 10000
```


# Computation

```{r}
set.seed(1)
data1 <- do(n_sample)*(predict(lm(y~x, data=generate_sample(f, n, sigma)), newdata=data.frame(x=c(x0))))

newdata <- generate_sample(f,n,sigma)
y_hat2 <- (predict(smooth.spline(x=newdata$x, y=newdata$y, df = 99), newdata=data.frame(x=.95)))

MSE1 <- mean((f(x0) - data1)^2)
MSE2 <- mean((f(x0) - y_hat2$y)^2)

variance1 <- var(data1)
variance2 <- var(y_hat2$y)

bias.sq1 <- ((f(x0) - mean(data1$X1))^2)
bias.sq2 <- ((f(x0) - mean(y_hat2$y))^2)

```


# Tables

As done in Lec 2.7, for both

* An `lm` regression AKA a `smooth.splines(x, y, df=2)` model fit 
* A `smooth.splines(x, y, df=99)` model fit 

output tables comparing:

|  MSE| bias_squared|   var| irreducible|   sum|
|----:|------------:|-----:|-----------:|-----:|
|     X|           X  |     X |      X |         X |

where `sum = bias_squared + var + irreducible`. You can created cleanly formatted tables like the one above by piping a data frame into `knitr::kable(digits=4)`.


```{r}
newdata <- data.frame(MSE= MSE1, bias_squared= bias.sq1, var= variance1, irreducible= MSE1 - (bias.sq1 + variance1), sum = bias.sq1 + variance1)
colnames(newdata) <- c("MSE", "bias_squared", "var", "irreducible", "sum")
head(knitr::kable(newdata, digits = 4))

newdata2 <- data.frame(MSE= MSE2, bias_squared= bias.sq2, var= variance2, irreducible= MSE2 - (bias.sq2 + variance2), sum = bias.sq2 + variance2)
colnames(newdata) <- c("MSE", "bias_squared", "var", "irreducible", "sum")
head(knitr::kable(newdata2, digits = 4))

```


# Analysis

**Questions**:

1. Based on the topics covered in Lec 2.7, name one possible "sanity check" for your results. Name another if you can.
1. In **two** sentences or less, give a rough sketch of what the procedure would
be to get the breakdown of $$\mbox{MSE}\left[\widehat{f}(x)\right]$$ for *all*
$x$ in this example, and not just for $$\mbox{MSE}\left[\widehat{f}(x_0)\right]
= \mbox{MSE}\left[\widehat{f}(0.95)\right]$$.
1. Which of the two models would you choose for predicting the point of interest and why?

**Answers**:

1. One possible sanity check for these results is to see if bias and variance are inversely related, as discussed in LEcture 2.7.

1. In order to do this, you would have to calculate the bias squared and viriance slightly differently, accounting for all the points in the model, as opposed to just the single point of interest.  Then, as we did for the single point of interest, you would add the values for bias squared and variance in order to get the MSE.

1. For predicting one point of interest, I would do the first method, where you predict using the linear model.  This is because in this method, there is less bias since the model is not overfit to the data as it is in the splines model where df=99.  There is more variance in the linerar model, but since the model isn't overfit to the data, it is a better model overall.